---
title: "Spatial Hedonic Model"
subtitle: "Subset (n=3000) of Lucas County housing data"
author: "Arnstein Gjestland"
institute: "HVL"
date: "Spring 2021 (revised autumn 2022)"
format:
  revealjs:
        theme: simple
        footer: '[MSB205](/index.html)'
        reveal_options:
        code-fold: show
        incremental: true
        smaller: true
        scrollable: true
        slide-number: c/t
editor: visual
editor_options:
  markdown:
    wrap: sentence
    canonical: true
    chunk_output_type: console
echo: true
eval: true
bibliography: [references.bib, LucasCounty-packages.bib]
nocite: '@*'
---

## Loading packages

```{r}
# To install urbnmapr from github
devtools::install_github("UrbanInstitute/urbnmapr")
```

```{r}
# for demean
suppressPackageStartupMessages({
library(Jmisc)
library(sf)
library(urbnmapr)
library(tidyverse)
library(ggplot2)
library(sp)
library(spdep)
library(spatialreg)
library(spData)
library(huxtable)
library(knitr)
})
# Options
options(scipen = 7)
```

```{r}
#| include: false
# create a bib file for the R packages
# used in this document
# Note! Needs to do a touch ag_model_basics.bib in terminal before first run
# else stops when bibliography: "ag_model_basics.bib" not found in YAML
knitr::write_bib(
  c(
  "Jmisc",
  "sf",
  "urbnmapr",
  "tidyverse",
  "ggplot2",
  "sp",
  "spdep",
  "spatialreg",
  "spData",
  "huxtable",
  "knitr"
    ),
  file = "LucasCounty-packages.bib"
  )
```

## Read in data

```{r}
data(house)
#head(house, n=10)
house_sf <- st_as_sf(house)
#head(house_sf, n=10)
dim(house_sf)
set.seed(42)# What else could it be?
red_house_sf <- house_sf[sample(1:dim(house_sf)[1], size = 3000), ]
```

## Quick check of data

```{r}
#| cache: true
plot(house_sf["price"], cex = 0.01, pch = 20)
```

```{r}
dim(red_house_sf)
```

## Dataset info (R help)

Data on 25,357 single family homes sold in Lucas County, Ohio, 1993-1998 from the county auditor.

Formal class 'SpatialPointsDataFrame' \[package "sp"\] with 5 slots.
The data slot is a data frame with 25357 observations on the following 24 variables.

price

:   a numeric vector

yrbuilt

:   a numeric vector

stories

:   a factor with levels one bilevel multilvl one+half two two+half three

TLA

:   a numeric vector

## Dataset info (R help)

wall

:   a factor with levels stucdrvt ccbtile metlvnyl brick stone wood partbrk

beds

:   a numeric vector

baths

:   a numeric vector

halfbaths

:   a numeric vector

frontage

:   a numeric vector

depth

:   a numeric vector

## Dataset info (R help)

garage

:   a factor with levels no garage basement attached detached carport

garagesqft

:   a numeric vector

rooms

:   a numeric vector

lotsize

:   a numeric vector

sdate

:   a numeric vector

avalue

:   a numeric vector

## Dataset info (R help)

s1993 to s1998

:   a numeric vector

syear

:   a factor with levels 1993 1994 1995 1996 1997 1998

age

:   a numeric vector

## Dataset info (R help)

Its projection is CRS(+init=epsg:2834), the Ohio North State Plane.

## Convert from SpatialPointsDataFrame to sf

We convert from SpatialPointsDataFrame to sf (simple feature, point).

```{r}
hist(100*house_sf$age, breaks = 30)
summary(100*house_sf$age)
```

```{r}
my_age <- as.numeric(levels(house_sf$syear))[as.integer(house_sf$syear)] - house_sf$yrbuilt
my_age <- my_age + 3
my_age1 <- 1999 - house_sf$yrbuilt
summary(100*house_sf$age - my_age)
hist(my_age1, breaks = 30)
summary(my_age1)
table(round(my_age1, 3) == round(100*house_sf$age, 3))
```

## Changing the age variable

Ok, the age variable is the age of the house 1999 (not when sold), divided by 100.
The age of the house when sold seems to be a more reasonable variable.

```{r}
# Trying to follow Bivand 2011, NHH Discussion Paper
# age seems to include strange values. Turns out to be age in 1999 divided by 100
# We calculate age at time of sale as age_sold = syear - yrbuilt
# syear is factor, convert to numeric
red_house_sf$age_sold <- as.numeric(levels(red_house_sf$syear))[as.integer(red_house_sf$syear)] - red_house_sf$yrbuilt
# Scale age_sold by dividing by 100
red_house_sf$age_sold <- red_house_sf$age_sold/100
hmod1 <- "I(log(price)) ~ I(age_sold) + I(age_sold^2) + 
I(age_sold^3) + I(log(TLA)) + I(log(lotsize)) + rooms + beds + syear"
```

## Running the first model

```{r}
lm1 <- lm(hmod1, data = red_house_sf)
```

```{r}
summary(lm1)
```

## Generate neighbour structure

**Find 5 nearest,** make neighbour list and finally convert the neighbour list to a row standardised weight matrix.

```{r}
red_house_sf_mat_nb <- knearneigh(red_house_sf, k = 5)
red_house_sf_nb <- knn2nb(red_house_sf_mat_nb)
red_house_sf_w <- nb2listw(red_house_sf_nb, style = "W")
```

Plot the price of a house against the mean price of its (5 nearest) neighbours.
There seems to be a positive spatial association

## Moran Plot

```{r}
moran.plot(log(red_house_sf$price), listw = red_house_sf_w, labels = FALSE, pch = 20, cex=0.3)
```

## Approximate profile-likelihood estimator (APLE)

Approximate profile-likelihood estimator (APLE), see [@OnestepEstimationSpatial]

```{r}
# Calculate APLE for the variables
red_house_sf %>% 
  select(price, age_sold, TLA, lotsize, rooms, beds) %>% 
  st_set_geometry(., NULL) %>% 
  demean() %>% 
  as_tibble() %>% 
  map(aple,listw = red_house_sf_w)
```

## Conclusion APLE

We see that the strength of positive spatial dependence is greatest for `price` and `lotsize` and lowest for number of bedrooms (`beds`) in the dwellings.
The slightly negative spatial association for the age of a sold house is perhaps a bit more surprising.

> \[APLE\], which may be viewed somewhat like a correlation coefficient between the values observed for each state and the average values for neighbouring states.[@bivandReviewSoftwareSpatial2021]

```{r}
# If we want to see which obs. are the most influential we can change labels to TRUE
# The row number for each influential obs. will then be plotted in the figure.
moran.plot(red_house_sf$age_sold, listw = red_house_sf_w, 
           labels = FALSE, pch = 20, cex=0.3)
```

```{r}
# Remember that age_sold is scaled by 1/100, so the max age 1.59 is in fact 159 years
summary(red_house_sf$age_sold)
```

## Spatial models

-   Local: SDEM, SEM, SLX, lm

-   Global: SDM, SAR, SLX, SEM, lm

## Local, start with SDEM

```{r SDEM, cache=TRUE}
# Some of these models are quite expensive to compute, so we cache the results
# It should automagically be recomputed if something change, but caching is complex
# so we should check that the models really are recomputed if we change the data or model
# Note the use of the option Durbin to choose the kind of model, used to be done by the options
# type and etype until recently.
SDEM <- errorsarlm(hmod1, data = red_house_sf, listw = red_house_sf_w, Durbin = TRUE)
```

```{r}
summary(impacts(SDEM), zstats = TRUE)
```

## Local, SEM

```{r SEM, cache = TRUE}
# No spatial effects in the explanatory variables
SEM <- errorsarlm(hmod1, data = red_house_sf, listw = red_house_sf_w, Durbin = FALSE)
```

```{r}
# normal summary, no impacts
summary(SEM)
```

## Local, SLX

```{r SLX}
# SLX is not costly to compute, so no caching
SLX <- lmSLX(hmod1, data = red_house_sf, listw = red_house_sf_w, Durbin = TRUE)
```

```{r}
# NEW
summary(SLX)
```

```{r}
summary(impacts(SLX))
```

## Choosing a local model

```{r}
LR.Sarlm(SEM, SDEM)
```

Discard H0, do not restrict to SEM.
We keep $\theta > 0$.

```{r}
LR.Sarlm(SDEM, SLX)
```

We discard H0 and will not restrict to SLX, continue with SDEM.

```{r}
LR.Sarlm(SDEM, lm1)
```

Discard H0, we will not restrict to OLS.

We end up with SDEM as the best local model (no endogeniety in the price variable).

```{r}
# residuals are with attributes, hence as.vector()
moran.plot(SDEM$residuals, listw = red_house_sf_w)
```

## Results

```{r}
# Do not use "SDEM" = SDEM, report coeffisent estimates not impacts
huxreg("OLS" = lm1, "SEM" = SEM, 
       error_format = "({statistic})",
       note = "{stars}. T-values reported.")
```

### Compare impacts from SDEM and SLX model

```{r}
# tidy(SDEM) or tidy(SLX) reports regression coefficients
# make our own table for easy comparison of impacts
i_SDEM <- impacts(SDEM)$impacts %>% 
  as_tibble() %>% 
  mutate(
    variable = names(impacts(SDEM)$impacts$direct)
  ) %>% 
  select(variable, everything())

i_SLX <- impacts(SLX)$impacts %>% 
  as_tibble()

names(i_SDEM)[2:4] <- paste("SDEM", names(i_SDEM)[2:4], sep = "_")
names(i_SLX) <- paste("SLX", names(i_SLX), sep = "_")
i_SDEM_SLX <- cbind(i_SDEM, i_SLX)
i_SDEM_SLX <- cbind(variable = i_SDEM_SLX[, "variable"], 
                    round(i_SDEM_SLX[, -1], 6))
kable(i_SDEM_SLX[, c(1:2,5,3,6,4,7)])
```

## Global

Spatial lag in y and X.

```{r SDM, cache=TRUE}
# expensive to compute
SDM <- lagsarlm(hmod1, data = red_house_sf, listw = red_house_sf_w, Durbin = TRUE)
```

```{r SDMimpacts, cache = TRUE}
# Computing the impacts is expensive too
impacts(SDM, listw = red_house_sf_w)
```

```{r SDMimpactsZ, cache = TRUE}
# Simulated, varies somewhat with each sample (note that seed is set above). Results might 
# differ with another seed
summary(impacts(SDM, listw = red_house_sf_w, R=500), zstats = TRUE)
```

```{r SAR, cache = TRUE}
SAR <- lagsarlm(hmod1, data = red_house_sf, listw = red_house_sf_w, Durbin = FALSE)
```

```{r}
# NEW
summary(SAR)
```

```{r SARimpactsZ, cache = TRUE}
summary(impacts(SAR, listw = red_house_sf_w, R=500), zstats = TRUE)
```

```{r}
LR.Sarlm(SDM, SAR)
```

We reject H0 and do not restrict to SAR.
Keep SDM.

```{r}
# SLX estimated above
LR.Sarlm(SDM, SLX)
```

We reject H0 and do not restrict to SLX.
Keep SDM.

```{r}
# Common factor?
LR.Sarlm(SDM, SEM)
```

We reject H0 and do not restrict to SEM.
Keep SDM.

```{r}
LR.Sarlm(SDM, lm1)
```

We reject H0 and do not restrict to OLS.
SDM seems to be the best global model.
We have spatial effects both in price and the explanatory variables.
There is a strong negative indirect effect regarding number of rooms that might be regarded as counter intuitive and should perhaps be looked further into.

```{r, cache=TRUE}
summary(impacts(SDM, listw = red_house_sf_w, R=1000), zstats = TRUE)
```

```{r}
# residuals are with attributes, hence as.vector()
moran.plot(as.vector(SDM$residuals), listw = red_house_sf_w)
```

## Some tests

## Are our residuals normal?

SDEM

```{r}
library(DescTools)
# H0: residuals normal
JarqueBeraTest(residuals(SDEM), robust = TRUE)
# Conclusion: not normal
```

Far from normal.

SDM

```{r}
# H0: residuals normal
JarqueBeraTest(residuals(SDM), robust = TRUE)
# Conclusion: not normal
```

Far from normal.

## Heteroskedasticity?

```{r}
# test for Heteroskedasticity
bptest.Sarlm(SDEM, studentize = TRUE)
```

```{r}
# test for Heteroskedasticity
bptest.Sarlm(SDM, studentize = TRUE)
```

## MAPS

```{r}
#  Start of code to draw a map of Lucas County
#
#
# counties %>% 
#   filter(county_name == "Lucas County")
# 
# 
# tmp <- countydata %>% 
#   left_join(counties, by = "county_fips") %>% 
#   filter(state_name =="Ohio") %>% 
#   filter(county_name == "Lucas County") %>% 
#   ggplot(mapping = aes(long, lat, group = group, fill = horate)) +
#   geom_polygon(color = "#ffffff", size = .25)
# 
# class(countydata)
# names(countydata)
# class(counties)
# names(counties)
# 
# tmp_sf <- st_as_sf(counties, coords = c("long", "lat"), 
#     crs = 2834)
# 
# names(tmp_sf)
# 
# tmp_sf %>% 
#   filter(state_name == "Ohio") %>% 
#   filter(county_name == "Lucas County") 
# 
# test <- st_cast(tmp_sf, "MULTIPOINT")
# test1 <- st_cast(test, "POLYGON")
# 
# names(tmp_sf)
# 
# tmp <- st_transform(tmp, 2834)
# tmp
# head(countydata)
```

```{r}
#siste
```
